Regression Cube Analysis of Cohort Study Data

### 2015-06-19 draft.tex
added reasoning on Support Design Decisions with Expert Feedback

While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
%%
\begin{enumerate}
	%%
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	%%
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
	%%
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
\com{TODO: Put into separate subsection?}
%%
\design{
%%
The \emph{Regression Cube} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
%%
The idea emerged from analysis sessions of a previous project, which contained a mosaic plot showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
%%
It allowed them to reproduce their knowledge by observing correlations they would expect as well as discovering new correlations.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature, often indicate the presence of the investigated disease.
%%
That is the focus of this work.
%%
% These target features often indicate the presence of the investigated disease.
%%
The domain experts wanted to \emph{model knowledge} about the investigated condition, such as confounding features (e.g., age or gender).
%%
For explorative analysis, they preferred an approach, which highlights associations w.r.t. various target features to both check for medical soundness of the data as well as detecting unexpected relationships.
%%
Additionally, due to the sensitive nature of population study data, the data has to be handled secure.
%%
This aspect is considered later in the Section~\ref{sec:SystemDesign}.
%%
%In this work, we want to provide an overview visualization of large cohort study data sets \emph{towards target features}.
}
%%
%The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
%Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
%%
%Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
%In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
%These target features often indicate the presence of the investigated disease.
%%

As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
%%
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
% \design{
% %%
% Key requirements based on the epidemiological application domain on the Regression Cube are:
% %%
% \begin{itemize}
% 	\item Show relationships of features w.r.t. their explanatory power towards the target feature (e.g., a disease or unhealthy condition).
% 	\item Allow for modeling assumptions regarding the target feature.
% 	\item The tool has to be easily accessible and simultaneously secure the data its processing.
% \end{itemize}
% %%
% }
\subsection{Cube Description Using Regression Formula Notation}
%%
Expert knowledge is integrated into the analysis using regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/cube}
 \caption{
 (a) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{and adjusted $R^2$} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{Regression Cube} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
%%
}

@@

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{Regression Cube} for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
Input formulas reflect \emph{hypotheses} about the data.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

\paragraph{Cube Comparison.}
%%
\emph{Regression Cubes} can be compared by creating difference views.
%%
One cube (extracted with one formula) acts as reference.
%%
The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
%%
For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)

\section{System Design} \label{sec:SystemDesign}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/system}
 \caption{
 %%
 Breast fat data set loaded into the \emph{Regression Cube} prototype.
 %%
 (a) Using the formula input, the user specifies the dependent feature as well as its calculation rules.
 %%
 (b) 3D cube visualization, showing values above the cube matrix diagonal as overview.
 %%
 The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
 %%
 (c) 2D mosaic plot visualization of the selected slice for feature \emph{Pain/Discomfort}.
 %%
 }
  \label{fig:System}
\end{figure}
%%
%Albeit our focus on epidemiological data sets, we designed our system to be open with open formats as input interfaces .
%%
We designed our system to be openly accessible and easy to use.
%%
With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
%%
The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.
%%
Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
%%
\\\\
%%
\design{
%A key requirement was little set up time to reduce the effort on the domain expert side, as they are mostly practicing clinicians with small time-frames for epidemiological research.
The design choice of using web-based technologies offers various advantages w.r.t. the collaboration with epidemiologists.
}
%Design choices and spaces are restricted by the underlying technologies and devices.
%%
%Therefore, the exchange of data and methods using web-based technologies offers many advantages.
%%
There is no set-up time involved and domain experts can use the methods from any computer connected with the web.
%%
By design, web technology is based on a client-server architecture, making it easy to outsource computationally heavy tasks on server clusters and transferring results to the client device.
%%
%The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
%%
%Therefore, we have to adapt our design respecting the conventions of web pages.
%%
\subsection{System Paradigm and Components}
%%
The \emph{Regression Cube} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
%%
This allows for a fast learning of the individual system parts.
%%
Our prototype consists of three components:
\begin{itemize}
	%%
	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
	%%
	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube [2.SI].
	%%
	\item The \emph{formula editor} allows formula input w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
	%%
	It also allows to select a reference formula for creating a difference cube [1.AF, 3.ZF].
	%%
\end{itemize}
%%
\paragraph{File Upload and Classification [1.AF].}
%%
Popular analytics tools, such as WEKA \cite{WEKA}, owe their popularity to their support of open file types.
%%
To allow other users even outside the epidemiological application domain to access to our tool, we use standard ASCII-based CSV files.
%%
The first line in a CSV file represents all features (columns) of the data set.
%%
Each line after that represents one subject (row) and its feature manifestations.

@@

\paragraph{Selecting and Scaling the Descriptive Regression Metric.}
\add{
%%
The descriptive metric shown in the 2D/3D view can be selected using a dropdown menu.
%%
The default selection is $R^2$.
%%
$AIC$ displays model quality.
%%
\emph{Adjusted} $R^2$ values are only available for linear regression, logistic regression results are represented via $R^2$ values in this mode.
%%
As they are visually distinguished using color, confusions are avoided.
%%
The transfer function of the color intensity (2D) and opacity (3D) can be adapted using a slider input.
%%
This additionally allows to filter models with desired features, such as only very high $R^2$ values.
}

\paragraph{3D Prism as Data Mini-Map.}
%%
The 3D cube representation acts as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
%%
It serves as a function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
%%
The displayed prism shows values above the matrix diagonal.
%%
For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System}b).
%%
Applying this strategy to a formula containing a static target (e.g. $Cancer \sim X + Y + Z$) yields many occlusions, since the CFS algorithm creates the same feature space for every slice.
%%
For such formulas, the 3D view encodes every data element as absolute difference between its $R^2$ features and the global mean along the z-axis.
%%
This highlights slices with unusually low or high results (Fig.~\ref{fig:Application}).
%%
\paragraph{Tackling the disadvantages of 3D information visualization.}
%%
3D information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
%%
We aim to minimize these problems.
%%
Since the $R^2$ values are mapped on data point opacity, large values are highlighted in the prism, guiding the focus to the respective slices.
%%
It also creates a sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
%%
Also, the preceding correlation-based feature selection reduces the information space significantly, leading to sparse cubes.
%%
Overlapping is still an issue, but this way greatly reduced in its effect to the readability of the visualization.
%%

>>>>

While the presented techniques were well received by the epidemiological experts, the explanatory power towards back pain was limited, yielding an analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
%These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is insufficient to characterize back pain.
\\\\
The difference of the presented approach towards our previous and related work is twofold:
%%
\begin{enumerate}
	%%
	\item We focus on the large-scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	%%
	\item The analysis is conducted w.r.t. a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
	%%
\end{enumerate}
%%
\section{Regression Cube Analysis of Cohort Study Data}
\design{
This section covers the design of the Regression Cube as well as how epidemiological requirements and  domain expert input was utilized to influence design decisions.
}
\subsection{Iterative Design Based on Expert Feedback}
%%
\design{
%%
The \emph{Regression Cube} design was developed iteratively based on feedback of epidemiologists by using the prototype in joint analysis sessions on their data sets.
%%
The idea emerged from analysis sessions of a previous project, which contained a mosaic plot showing pairwise feature correlations based on \emph{Cram\'{e}r's V} contingency values \cite{Klemm2014VIS}.
%%
It allowed them to reproduce their knowledge by observing correlations they would expect as well as discovering new correlations.
%%
In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature, often indicate the presence of the investigated disease.
%%
That is the focus of this work.
%%
% These target features often indicate the presence of the investigated disease.
%%
The domain experts wanted to \emph{model knowledge} about the investigated condition, such as confounding features (e.g., age or gender).
%%
For explorative analysis, they preferred an approach, which highlights associations w.r.t. various target features to both check for medical soundness of the data as well as detecting unexpected relationships.
%%
Additionally, due to the sensitive nature of population study data, the data has to be handled secure.
%%
This aspect is considered later in the Section~\ref{sec:SystemDesign}.
%%
%In this work, we want to provide an overview visualization of large cohort study data sets \emph{towards target features}.
}
%%
%The basic idea of our \emph{Regression Cube} is to provide an overview visualization of large cohort study data sets towards target features.
%%
%Overview visualizations of feature relationships as presented by Angelelli et al.~\cite{Angelelli} are often focused on relationships between the visualized features.
%%
%Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
%In epidemiology, these relationships are also of interest, but rather w.r.t. their explanatory power towards the target feature.
%%
%These target features often indicate the presence of the investigated disease.
%%

As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
%%
A regression model is based on expert knowledge; there is no singular rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
% \design{
% %%
% Key requirements based on the epidemiological application domain on the Regression Cube are:
% %%
% \begin{itemize}
% 	\item Show relationships of features w.r.t. their explanatory power towards the target feature (e.g., a disease or unhealthy condition).
% 	\item Allow for modeling assumptions regarding the target feature.
% 	\item The tool has to be easily accessible and simultaneously secure the data its processing.
% \end{itemize}
% %%
% }
\subsection{Cube Description Using Regression Formula Notation}
%%
\design{The key requirement of modeling expert knowledge} is integrated into the analysis using \emph{regression formulas}.
%Expert knowledge is integrated into the analysis using regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%

@@

%\subsection{Abstracting Regression Results Using $R^2$}
\subsection{Abstracting Regression Results}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/cube}
 \caption{
 (a) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $age$.
 %%
 The descriptive metrics extracted from the regression model are mapped to color saturation (for $R^2$, saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The metric describing the regression model of each slice voxel is mapped on opacity in the 3D view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
\noindent The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%%
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ \add{and adjusted $R^2$} value are metrics allowing for this kind of assessment.
%%
% \add{
% %%
% Users with in-depth knowledge about regression models are given the option to render the \emph{Regression Cube} using other descriptive features presented in \ref{sec:RegressionAnalysis}.
% %%
% Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
% %%
% To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
% }
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between $[0, 1]$, it allows for comparison \emph{between} regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube}a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%
\design{
%%
Based on expert feedback on early versions of this view, the amount of features used to compare regression models was extended.
%%
Therefore, these experts can investigate the cube with emphasis on specific aspects of the model.
%%
}
%%
\add{
%%
$Adjusted$ $R^2$ can be represented in the same way, since they share the scale.
%%
$AIC$ values are not restricted to a normalized scale.
%%
They have to be normalized in order to map it onto color saturation.
%%
%The resulting scale may be distort by outliers, which requires facilities for allowing the user to select scale ranges.
The resulting scale may be distorted by outliers derived from poor regression models.
%%
To tackle this problem, we provide a slider input, which maps the transfer function of the metric to color saturation based on user-selected ranges.
%%
Outliers can be cut off to emphasize ranges of interest.
%%
}
\design{
%%
To avoid shutting out users unfamiliar with these metrics, the \emph{Regression Cube} is set per default to show $R^2$ values.
%%
}

@@

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation to the system design section later on.
%%
As shown in Fig.~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or to create a new \emph{Regression Cube} for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
\design{
%%
Early analysis sessions yielded two approaches on analyzing the data.
%%
The classic approach is the \emph{hypothesis-based} approach, where the expert already has a good understanding of the data and potential associations (e.g. reproducing knowledge about hepatic steatosis risk factors based on known risk factors).
%%
The \emph{hypothesis-free} approach is well suited for deriving new insight into the data, such as identifying cofounding features or potential targets (e.g. deriving risk factors for breast cancer associated features).
%%
}
%%
%Input formulas reflect \emph{hypotheses} about the data.
\emph{Hypotheses} about the data are reflected using input formulas.
%%
Using the operators, dynamic variables and dataset features, many different assumptions can be expressed.
%%
To support the \emph{hypothesis-free} analysis, we provide a default formula:

@@

\paragraph{Cube Comparison.}
\design{
%%
Cube comparisons were introduced later in the project, as epidemiologists with focus on statistics pointed out, that comparing outcomes of different formulas is suitable for removing the effect of possible confounding features.
%%
}
%%
\emph{Regression Cubes} can be compared by creating difference views.
%%
One cube (extracted with one formula) acts as reference.
%%
The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
%%
For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$).

\section{System Design} \label{sec:SystemDesign}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/system}
 \caption{
 Breast fat data set loaded into the \emph{Regression Cube} prototype.
 (a) Using the formula input, the user specifies the dependent feature as well as its calculation rules.
 (b) 3D cube visualization, showing values above the cube matrix diagonal as overview.
 The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
 (c) 2D mosaic plot visualization of the selected slice for feature \emph{Pain/Discomfort}.
 }
  \label{fig:System}
\end{figure}
%Albeit our focus on epidemiological data sets, we designed our system to be open with open formats as input interfaces .
We designed our system to be openly accessible and easy to use.
With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses, maybe with other (statistical) tools.
Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
\\\\
\design{
%A key requirement was little set up time to reduce the effort on the domain expert side, as they are mostly practicing clinicians with small time-frames for epidemiological research.
The design choice of using web-based technologies offers various advantages w.r.t. the collaboration with epidemiologists.
Since they are usually practicing clinicians, they usually have little time to wrangle software.
A web-based approach has no set-up time besides loading up the data set and can be carried out with any computer connected with the web.
Feedback of domain experts even can be implemented directly in analysis sessions and just requires a page reload.
By providing a service a website it has a much larger chance of being tested and potentially adapted by a wide range of experts.
}
%Design choices and spaces are restricted by the underlying technologies and devices.
%Therefore, the exchange of data and methods using web-based technologies offers many advantages.
%There is no set-up time involved and domain experts can use the methods from any computer connected with the web.
By design, web technology is based on a client-server architecture, making it easy to outsource computationally heavy tasks on server clusters and transferring results to the client device.
%The design space spanned by web technologies is different to standard WIMP applications: right-click menus, modal windows and menu bars are no established user interface components in this context.
%Therefore, we have to adapt our design respecting the conventions of web pages.
\subsection{System Paradigm and Components}
\design{
Complex system, which take a lot of training and experience to work with are not suitable for the task of deriving overview visualizations for epidemiologists, as they have little time slots available for their research.
If the system is not easy to learn, it will not be adapted.
}
%The \emph{Regression Cube} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
Therefore, the \emph{Regression Cube} design focuses on a clean appearance, reducing the amount of user-interface elements as much as possible.
This allows for a fast learning of the individual system parts.
Our prototype consists of three components:
	\item The \emph{file upload} section starting the analysis with providing a comma-separated value (CSV) file [1.AF].
	\item The \emph{cube visualization} consisting of the 2D mosaic view as well as a 3D representation of the whole cube \add{with facilities to change the represented regression metric and its range} [2.SI].
	\item The \emph{formula editor} allows formula input w.r.t. a hypothesis or to conduct a \emph{hypothesis-free} analysis.
	It also allows to select a reference formula for creating a difference cube [1.AF, 3.ZF].
\paragraph{File Upload and Classification [1.AF].}
Popular analytics tools, such as WEKA \cite{WEKA}, owe part of their success to their support of open file types.
To allow other users even outside the epidemiological application domain to access to our tool, we use standard ASCII-based CSV files.
The first line in a CSV file represents all features (columns) of the data set.
Each line after that represents one subject (row) and its feature manifestations.

@@

\paragraph{Selecting and Scaling the Descriptive Regression Metric.}
\design{
%%
As the feedback made apparent that other features are of interest for analyzing regression models too, UI elements for controlling them were introduced to the system.
%%
}
\add{
%%
The descriptive metric shown in the 2D/3D view can be selected using a dropdown menu.
%%
The default selection is $R^2$.
%%
$AIC$ displays model quality.
%%
\emph{Adjusted} $R^2$ values are only available for linear regression, logistic regression results are represented via $R^2$ values in this mode.
%%
As they are visually distinguished using color, confusions are avoided.
%%
The transfer function of the color intensity (2D) and opacity (3D) can be adapted using a slider input.
%%
This additionally allows to filter models with desired features, such as only very high $R^2$ values.
}

\paragraph{3D Prism as Data Mini-Map.}
\design{
%%
In early prototype versions, the 3D prism acted as starting point for the data analysis without the implementation of a separate 2D view.
%%
Slices were shown using cutaway planes.
%%
This approach was not popular with our domain experts, because the complexity of the visualization overwhelmed them. %did not like the combination of 3D interaction and slicing with representatives.
%%
}
%%
The 3D cube representation \design{was redesigned to} act as an overview over the whole data set, but its purpose is not to derive detailed information about data points.
%%
It serves as a function similar to a mini-map, guiding the attention towards points of interest in the data, as well as giving context information about adjacent data values when using the 2D mosaic plot.
%%
\design{The distinction between overview and details and demand using two different representations was well received with our domain experts.}
%%
The displayed prism shows values above the matrix diagonal.
%%
For formulas with a dynamic target feature (e.g. exploratory analysis using $Z \sim X + Y$), the color encodes absolute $R^2$ values (Fig.~\ref{fig:System}b).
%%
Applying this strategy to a formula containing a static target (e.g. $Cancer \sim X + Y + Z$) yields many occlusions, since the CFS algorithm creates the same feature space for every slice.
%%
For such formulas, the 3D view encodes every data element as absolute difference between its $R^2$ features and the global mean along the z-axis.
%%
This highlights slices with unusually low or high results (Fig.~\ref{fig:Application}).
%%
\paragraph{Tackling the disadvantages of 3D information visualization.}
%%
3D information visualizations are often criticized for introducing occlusions and interaction problems, which often do not balance out the advantages of using the third dimension for visual mapping.
%%
We aim to minimize these problems.
%%
Since the $R^2$ values are mapped on data point opacity, large values are highlighted in the prism, guiding the focus to the respective slices.
%%
It also creates a sparse representation, since the majority of regression models yield (depending on the data set and the chosen formula) low $R^2$ values.
%%
Also, the preceding correlation-based feature selection reduces the information space significantly, leading to sparse cubes.
%%
Overlapping is still an issue, but this way greatly reduced in its effect to the readability of the visualization.
%%
